---
title: "simulate_alien_data"
author: "Martine Lind Jensen"
date: "2024-04-26"
output: html_document
---
We have to simulate data from the experiment 

```{r}
pacman::p_load(tidyverse, cmdstanr, future, furrr, brms)
```

```{r}
df <- read_delim("AlienData.txt")
```

Structure of the experiment:

- the stimuli are conceptualized as 5 dimensional vectors of 0s and 1s (5 features, binary values)

- there are 32 possible stimuli, all 32 stimuli are presented in randomized order, in three iterations (stimuli 1-32 in random order, stimuli 1-32 in a new random order, stimuli 1-32 in a new random order).

- the stimuli are categorized along two dimensions: danger (0-1) and nutrition (0-1). Feel free to simplify your life and only consider one dimension (but kudos for considering both).
the association between feature and category varies over session. 

- In the first session: danger depends on the alien having spots AND eyes on stalks (feature 1 AND feature 2 both being 1); nutrition depends on arms being up (feature 4 being 1).
    **We only need to model first session**


Simulating the data and full sequence of stimuli shown in the experiment. We only work with dangerous or not, but we include all 5 features of the feature space.

**PLOT NEEDS CLEAN-UP**

```{r}
#creating stimuli, each feature can be 0 or 1 (up or down, blue or green, on or off)
arms <- c(0,1)
legs <- c(0,1)
eyes <- c(0,1) #eyes on 1 = dangerous
spots <- c(0,1) #spots on 1 = dangerous
color <- c(0,1)

#Expanding the feature grid to have the 32 stimuli
feature_grid <- expand.grid(arms = arms, legs = legs, eyes = eyes, spots = spots, color = color)

#Selecting which stimuli is dangerous 
feature_grid <- feature_grid %>% mutate(
  dangerous = ifelse(spots == 1 & eyes == 1, dangerous <- 1, dangerous <- 0)
)

#shuffle the stimuli 
experiment <- feature_grid[sample(nrow(feature_grid)), ]

#Plotting the 32 stimuli (not done)
ggplot(stimuli) + 
  geom_point(aes(arms, dangerous), color = "red", position = "jitter") +
  geom_point(aes(legs, dangerous), color = "green", position = "jitter") +
  geom_point(aes(spots, dangerous), color = "blue", position = "jitter") +
  geom_point(aes(eyes, dangerous), color = "orange", position = "jitter") +
  geom_point(aes(color, dangerous), color = "yellow", position = "jitter") +
  facet_wrap(.~ row.names(stimuli))

```



CODE FROM LECTURE: to see the structure used  
```{r CODE FROM LECTURE}
# Defining the stimuli, their height and position features, and their category
stimulus1 <- c(5,3,7,1,8,2,6,4)
height <- c(1,1, 2,2, 3,3, 4,4)
position <- c(2,3, 1,4, 1,4, 2,3)
category <- as.factor(c(0,0, 1,0, 1,0, 1,1))

# Making this into a tibble
stimuli1 <- tibble(stimulus1, height, position, category)



# Generating the sequence of stimuli in the full experiment
sequence <- c()
for (i in 1:8) {
  temp <- sample(seq(8), 8, replace = F)
  sequence <- append(sequence, temp)
}

experiment1 <- tibble(stimulus1 = sequence, height = NA, position = NA, category = NA)
for (i in seq(nrow(experiment1))) {
  experiment1$height[i] <- stimuli1$height[stimuli1$stimulus1 == experiment1$stimulus1[i]]
  experiment1$position[i] <- stimuli1$position[stimuli1$stimulus1 == experiment1$stimulus1[i]]
  experiment1$category[i] <- as.numeric(as.character(stimuli1$category[stimuli1$stimulus1 == experiment1$stimulus1[i]]))
}
```

```{r distance and similarity functions}
# Distance 
distance <- function(fact1, fact2, weight) {
  return(sum(weight * abs(fact1 - fact2)))
}

# Similarity
similarity <- function(distance, c) {
  return(exp(-c * distance))
}

# Let's assess similarity
dd <- tibble(
  expand_grid(
    distance = c(0,.1,.2, .3,.4,.5,1,1.5,2,3,4,5,6), 
    c = c(0.1, 0.2, 0.5, 0.7, 1, 1.5, 2, 3, 4, 5, 6))) %>% 
  mutate(
    similarity = similarity(distance, c)
  )

#Plotting
dd %>% mutate(c = factor(c)) %>%
  ggplot() +
  geom_line(aes(distance, similarity, group = c, color = c)) + 
  theme_bw()
```

```{r gcm function}
### generative model ###
gcm <- function(weight, c, obs, cat_dangerous) {
  
  # create an empty list to save probability of saying "1" for each trial
  r <- c()
  
  ntrials <- nrow(obs)
  
  for (i in 1:ntrials) {

    # if this is the first trial, or there any category with no exemplars seen yet, set the choice to random
    if (i == 1 || sum(cat_dangerous[1:(i - 1)]) == 0 || sum(cat_dangerous[1:(i - 1)]) == (i - 1)) {
      r <- c(r, .5)
    } else {
      similarities <- c()
      # for each previously seen stimulus assess distance and similarity
      for (e in 1:(i - 1)) {
        sim <- similarity(distance(obs[i, ], obs[e, ], weight), c)
        similarities <- c(similarities, sim)
      }
      # Calculate prob of saying "1" by dividing similarity to 1 by the sum of similarity to 1 and to 2
      numerator <- 0.5 * sum(similarities[cat_dangerous[1:(i - 1)] == 1])
      denominator <- 0.5 * sum(similarities[cat_dangerous[1:(i - 1)] == 1]) + 0.5 * sum(similarities[cat_dangerous[1:(i - 1)] == 0])
      r <- c(r, numerator / denominator)
    }
  }

  return(rbinom(ntrials, 1, r))
}
```

```{r simulated response function}
# function for simulation responses
simulate_responses <- function(agent, weight, c) {
    
    observations <- experiment %>%
        select(c("arms", "legs", "spots", "eyes", "color"))
    
    dangerous <- experiment$dangerous
    
    if (weight == "equally") {
        weighted <- rep(0.2,5) #weighting each feature equally with 0.2 
    } else if (weight == "optimally") {
        weighted <- c(0, 0, 0.5, 0.5, 0) #Weighting spots and eyes equally since they are the dangerous features. Given the other features a weight of 0
    } 

    # simulate responses
    responses <- gcm(
        weighted,
        c,
        observations,
        dangerous
    )
    
    tmp_simulated_responses <- experiment %>%
        mutate(
            trial = seq(nrow(experiment)),
            sim_response = responses,
            correct = ifelse(dangerous == sim_response, 1, 0),
            performance = cumsum(correct) / seq_along(correct),
            c = c,
            weight = weight,
            agent = agent
        )

    return(tmp_simulated_responses)
}


# simulate responses
plan(multisession, workers = availableCores())

param_df <- dplyr::tibble(
    expand_grid(
        agent = 1:10,
        c = seq(.1, 2, 0.2),
        weight = c("equally", "optimally")
    )
)

simulated_responses <- future_pmap_dfr(param_df,
    simulate_responses,
    .options = furrr_options(seed = TRUE)
)
```


```{r plotting simulated responses}

#Optimally by agent
simulated_responses %>% 
  subset(weight == "optimally") %>% 
  ggplot(aes(x = trial, y = performance, color = as.factor(agent))) +
  geom_line () +
  facet_wrap(.~ c) +
  labs(x = "Trial", y = "Performance", color = "Agent") + 
  theme_minimal()+ 
  ggtitle("Weighted optimally")

#Equally by agent
simulated_responses %>% 
  subset(weight == "equally") %>% 
  ggplot(aes(x = trial, y = performance, color = as.factor(agent))) +
  geom_line () +
  facet_wrap(.~ c) +
  labs(x = "Trial", y = "Performance", color = "Agent") + 
  theme_minimal()+ 
  ggtitle("Weighted equally")

#Averaged over agent 
ggplot(simulated_responses, aes(trial, performance, group = weight, color = weight)) +
  geom_smooth() +
  theme_minimal() +
  labs(x = "Trial", y = "Performance", color = "Weight") + 
  facet_wrap(c ~ .)

ggplot(simulated_responses, aes(trial, performance, group = as.factor(c), color = as.factor(c))) +
  geom_smooth() +
  theme_minimal() +
  labs(x = "Trial", y = "Performance", color = "C") + 
  facet_wrap(weight ~ .)

```